---
title: "Corrected Covered Area (CCA) in Systematic Reviews"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Corrected Covered Area (CCA)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction

The **Corrected Covered Area (CCA)** quantifies the overlap of studies included in multiple systematic reviews. A high CCA value indicates a high degree of overlap, suggesting potential redundancy in evidence synthesis.

This vignette explains:

- The formula for CCA

- How to compute CCA in R

- How to visualize study overlap using a heatmap

## CCA Formula

CCA is calculated as:

\[
CCA = \frac{N - r}{r (c - 1)} \times 100
\]

Where:
- \( N \) = total number of study occurrences across reviews
- \( r \) = number of unique included studies
- \( c \) = number of systematic reviews



## Example Data

We use an example dataset with **5 studies** across **5 systematic reviews**:

```{r, echo=TRUE, warning=FALSE, message=FALSE}
library(tidyverse)
library(junoR)  

# Example dataset
data <- tibble(
  study = rep(paste0("Study_", 1:5), times = 5),
  review = rep(paste0("Review_", 1:5), each = 5),
  status = c("included", "included", "ineligible", NA, NA, 
             NA, "included", NA, "excluded", NA, 
             "included", NA, "included", "excluded", "ineligible", 
             "included", "excluded", "included", "excluded", NA, 
             NA, "included", "ineligible", NA, "excluded")
)

# View the data
print(data)
```

## Compute CCA

We use the `cca_heatmap()` function to compute CCA:

```{r,echo=TRUE, warning=FALSE, message=FALSE}
# Run the function
result <- cca_heatmap(data)

# Print CCA result
result$CCA

```

This means that `r result$CCA`% of the included studies are redundant.

## Visualising Study Overlap

The function also generates a heatmap showing which studies are included in which reviews:

```{r}
# Generate heatmap
result$plot
```


This produces a heatmap where:

Green = included studies

Dark grey = excluded studies

Orange = ineligible studies

Light grey = missing data

## Interpretation of Results

|CCA (%)|	Overlap Level|	Interpretation|
|-------|--------------|----------------|
|0-20%|	Low|	Minimal study redundancy across reviews|
|21-50%|	Moderate|	Some overlap, but different studies are included|
|51-100%|	High|	High redundancy, reviews rely on similar studies|

>100%	Extreme	Indicates a possible error in study categorization


## References

Pieper, D., Antoine, S. L., Mathes, T., Neugebauer, E. A., & Eikermann, M. (2014).
Systematic review finds overlapping reviews were not mentioned in every other overview.
Journal of Clinical Epidemiology, 67(4), 368–375. DOI: 10.1016/j.jclinepi.2013.11.007

Hedges, L. V., & Olkin, I. (1988). Statistical Methods for Meta-Analysis. Academic Press.
