---
title: "An example seeded topic model"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{An example seeded topic model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# 1. What is LDA?
Latent Dirichlet Allocation (LDA) is a probabilistic topic modelling algorithm used to discover hidden topics in a collection of documents.

- Each document is represented as a mixture of topics.

- Each topic is represented as a mixture of words.

`Example: If you analyse news articles, LDA might find topics like "Politics," "Sports," or "Technology," each characterised by related words.`

# 2. What is Seeded LDA?

## Problem with Standard LDA

LDA is an unsupervised method—it doesn't know what topics should be, so the discovered topics may not always align with what you expect.

## Solution: Seeded LDA

In Seeded LDA, we provide seed words for each topic.
This helps guide the topic formation process so that topics align with meaningful categories.

**Example Use Case:**

If analyzing scientific literature, we may seed topics with specific words:

"Climate Change" → ["warming", "carbon", "emissions"]

"Biodiversity" → ["species", "habitat", "conservation"]

# 3. Seeded LDA Example in R
We will use the textmineR package to perform Seeded LDA.

## Step 1: Install and Load Packages

```{r, warning=FALSE, message=FALSE}
#install.packages("textmineR")
#install.packages("tm")  # For text preprocessing

library(textmineR)
library(tm)

```

## Step 2: Load Example Text Data

We will use a small dataset of articles.

```{r, warning=FALSE, message=FALSE}
documents <- c(
  "Climate change is causing rising temperatures and increasing emissions.",
  "Renewable energy sources like solar and wind help reduce carbon footprint.",
  "Biodiversity is under threat due to habitat destruction and pollution.",
  "Technology companies invest in artificial intelligence and data science.",
  "Deep learning and neural networks are advancing AI research."
)
```

Each document is a separate text entry, and we want to identify themes in them.

## Step 3: Preprocess Text Data

LDA works best with cleaned text. We will:

* Convert text to lowercase
* Remove stopwords (e.g., "the", "is", "and")
* Remove punctuation
* Convert text into a document-term matrix (DTM)

```{r, warning=FALSE, message=FALSE}
# Create a corpus
corpus <- Corpus(VectorSource(documents))

# Preprocess: Convert to lowercase, remove stopwords, remove punctuation
corpus <- tm_map(corpus, content_transformer(tolower))
corpus <- tm_map(corpus, removeWords, stopwords("english"))
corpus <- tm_map(corpus, removePunctuation)

# Create a document-term matrix (DTM)
dtm <- CreateDtm(doc_vec = sapply(corpus, as.character), 
                 doc_names = paste0("doc", 1:length(documents)), 
                 ngram_window = c(1,2))  # Unigrams & bigrams

```

## Step 4: Define Seed Words

We guide the model using seed words for each topic:

```{r, warning=FALSE, message=FALSE}
# Define seed words for topics
seed_words <- list(
  "climate_change" = c("climate", "warming", "carbon", "emissions"),
  "biodiversity" = c("biodiversity", "species", "habitat", "pollution"),
  "technology" = c("technology", "artificial", "intelligence", "data")
)

```

### Why Seed Words?

This ensures that LDA associates documents with topics based on expert knowledge.
Otherwise, random words might define topics in unexpected ways.

## Step 5: Fit Seeded LDA Model

We train the Seeded LDA model:

```{r, warning=FALSE, message=FALSE}
# Train Seeded LDA
lda_model <- FitLdaModel(dtm = dtm, 
                         k = length(seed_words),  # Number of topics
                         iterations = 500, 
                         burnin = 100,
                         seed_words = seed_words)
```

### Parameters Explained
* k = length(seed_words): Number of topics (based on how many seed word groups we define).

* iterations = 500: Number of updates to the model.

* burnin = 100: Initial iterations ignored for better convergence.

## Step 6: View Topic Assignments

After training, check which topics dominate each document:

```{r, warning=FALSE, message=FALSE}
lda_model$theta  # Topic proportions per document
```
Here each row = a document and each column = a topic. 

# 4. How to Interpret Results?

Each topic is a probability distribution over words.
Each document is a probability distribution over topics.
Words that strongly define topics can be extracted:

```{r}
get_top_terms <- function(phi, num_words = 10) {
  apply(phi, 1, function(topic) {
    names(sort(topic, decreasing = TRUE))[1:num_words]
  })
}

get_top_terms(lda_model$phi)  # Top words for each topic
```

# 5. Standard LDA vs. Seeded LDA

Feature	|Standard LDA|	Seeded LDA
Topic Discovery	|Unsupervised (random topics)|	Semi-supervised (guided topics)
Accuracy|	Can be noisy|	More aligned with domain knowledge
Flexibility|	Fully data-driven|	Requires predefined seed words
Use Case|	General topic discovery|	When you have expected topics

## When to Use Seeded LDA?

* You expect specific topics and want to guide LDA results.

* You don't want topics to be random.









