---
title: "Using R to clean up text"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using R to clean up text}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

When working with text data imported from, for example Excel, you may encounter issues such as:

- Non-printable characters (e.g., zero-width spaces, control characters)
- Carriage returns (`\r`) and line breaks (`\n`)
- Bullet points and other unwanted symbols

This document outlines common techniques for cleaning such text using R.

## Loading Required Packages

```{r setup, message=FALSE, warning=FALSE}
library(stringr)
```

## Removing Non-Printable Characters

Non-printable characters, such as ASCII control characters, can be removed using `iconv()` or `str_replace_all()`.

```{r}
clean_nonprintable <- function(text) {
  text <- iconv(text, "latin1", "ASCII", sub = "") # Removes non-ASCII characters
  text <- str_replace_all(text, "[^[:print:]]", "") # Removes other control characters
  return(text)
}

# Example
test_text <- "Hello\u0007World!"
clean_nonprintable(test_text)
```

## Removing Carriage Returns and Line Breaks

Excel often includes carriage returns (`\r`) and new lines (`\n`) in text data. These can be removed using `gsub()` or `str_replace_all()`.

```{r}
clean_linebreaks <- function(text) {
  text <- str_replace_all(text, "\r|\n", " ") # Replace with space
  return(text)
}

# Example
test_text <- "This is line one.\r\nThis is line two."
clean_linebreaks(test_text)
```

## Removing Bullets and Special Symbols

Bullet points and other special characters can be removed using regular expressions.

```{r}
clean_bullets <- function(text) {
  text <- str_replace_all(text, "[•●▪▶✓✔]", "") # Remove common bullet points
  return(text)
}

# Example
test_text <- "• Item one\n● Item two"
clean_bullets(test_text)
```

## Applying All Cleaning Functions

You can create a single function that applies all these cleaning steps.

```{r}
clean_text <- function(text) {
  text <- clean_nonprintable(text)
  text <- clean_linebreaks(text)
  text <- clean_bullets(text)
  text <- str_trim(text) # Trim leading/trailing spaces
  return(text)
}

# Example
test_text <- "• Item one\r\n● Item two with \u0007non-printable chars"
clean_text(test_text)
```

## Conclusion

These functions help standardise imported text, making it easier to process and analyse in R. 

